\texttt{Supertree construction with}\texttt{\emph{ Incertae sedis}}\texttt{ taxa}

### 2018-03-27 doc/incertae_sedis_latex2.tex
Update ms.

After solving a supertree (sub) problem, we need to assign taxon names
to the supertree nodes based on the taxonomy tree in the problem.
Each taxon name $n$ corresponds to a split $S(n)=S(n)_{1}|S(n)_{2}$
on the corresponding branch of the taxonomy tree. Without \emph{incertae
sedis}, such splits are always of the form $S(n)_{1}|\mathcal{L}-S(n)_{1}$,
but with \emph{incertae sedis} taxa $S_{2}(n)$ may be smaller than
$\mathcal{L}-S(n)_{1}$.

@@

However, in the \emph{incertae sedis} framework, it is possible for
one name to apply to multiple nodes. For example, in Figure \ref{fig:One-name-can},
the name $A$ can apply to multiple nodes. Here the name $A$ corresponds
to the split $A1\,A2\,|\bullet B\,C$, leaving out $D$ since it is
\emph{incertae sedis}. The two nodes $x$ and $y$ imply the splits
$A1\,A2\,|\bullet B\,C\,D$ and $A1\,A2\,D|\bullet B\,C$ respectively,
and both of these splits imply the split $A1\,A2\,|\bullet B\,C$,
so the name $A$ can apply to both $x$ and \footnote{Define ``implies'', as in the propinquity paper: $A_{1}|\bullet A_{2}\implies B_{1}|\bullet B_{2}$
means $B_{1}\subseteq A_{1}$ and $B_{2}\subseteq A_{2}$.}. This cannot happen without\emph{ incertae sedis} taxa except at
monotypic nodes. When faced with a choice about where to place a name,
we find the most tip-ward node where the name can apply and attach
the name to this node. This corresponds to a conservative choice about
whether \emph{incertae sedis} taxa should be added to an existing
taxon.
\begin{figure}
\hfill{}\subfloat[Taxonomy tree]{\includegraphics[scale=0.5]{Figures/name1nodes2/tax-shaded}

@@

It is also possible for multiple names to apply to a single node.

>>>>

After solving a supertree (sub) problem, we need to assign taxon names
to the supertree nodes based on the taxonomy tree in the problem.
Each taxon name $n$ corresponds to a split $S(n)=S(n)_{1}|\bullet S(n)_{2}$
on the corresponding branch of the taxonomy tree. Without \emph{incertae
sedis}, such splits are always of the form $S(n)_{1}|\bullet\mathcal{L}-S(n)_{1}$,
but with \emph{incertae sedis} taxa $S_{2}(n)$ may be smaller than
$\mathcal{L}-S(n)_{1}$.

@@

However, in the \emph{incertae sedis} framework, it is possible for
one name to apply to multiple nodes. For example, in Figure \ref{fig:One-name-can}b,
the name $A$ can apply to nodes $x$ and $y$. Here the name $A$
corresponds to the split $A1\,A2\,|\bullet B\,C$, leaving out $D$
since it is \emph{incertae sedis}. The two nodes $x$ and $y$ imply
the splits $A1\,A2\,|\bullet B\,C\,D$ and $A1\,A2\,D|\bullet B\,C$
respectively, and both of these splits imply the split $A1\,A2\,|\bullet B\,C$,
so the name $A$ can apply to both $x$ and $y$\footnote{Define ``implies'', as in the propinquity paper: $A_{1}|\bullet A_{2}\implies B_{1}|\bullet B_{2}$
means $B_{1}\subseteq A_{1}$ and $B_{2}\subseteq A_{2}$.}. This cannot happen without\emph{ incertae sedis} taxa except at
monotypic nodes. When faced with a choice about where to place a name,
our solution is to find the most tip-ward node where the name can
apply and attach the name to this node. This corresponds to a conservative
choice about whether \emph{incertae sedis} taxa should be added to
an existing taxon.
\begin{figure}
\hfill{}\subfloat[Taxonomy tree]{\includegraphics[scale=0.5]{Figures/name1nodes2/tax-shaded}

@@

It is also possible for multiple names to apply to a single node.
For example, in Figure \ref{fig:Multiple-names-can}a, the taxon $A$
corresponds to the split $B1\,B2\,C\,|\bullet\,Y$, and is child taxon
$B$ corresponds to the split $B1\,B2\,|\bullet\,Y$. The edge leading
to $A$ is consistent with the split for $B$, but the name $B$ is
applied to the node with the smallest include group. However, in the
solution tree (Fig. \ref{fig:Multiple-names-can}b), there is is only
node for both names $A$ and $B$ to apply to. In this case, we solve
this problem by introducing a monotypic parent, and applying $A$
to the newly created parent node (Fig. \ref{fig:Multiple-names-can}c).
\begin{figure}
\hfill{}\subfloat[Taxonomy tree]{\includegraphics[scale=0.5]{Figures/names2nodes1/tax-shaded}

}\hfill{}\subfloat[Solution tree]{\includegraphics[scale=0.5]{Figures/names2nodes1/synth-shaded}

}\hfill{}\subfloat[Solution tree]{\includegraphics[scale=0.5]{Figures/names2nodes1/solution2-shaded}

\caption{\label{fig:Multiple-names-can}Multiple names can apply to a single
node.}

Not all cases where multiple names apply to a single node can be resolved
in this way.
